### 2017-12-03 reports/abstract/project-abstract.tex
Clean up repo.

%        File: project-proposal.tex
%     Created: Wed Oct 11 11:00 PM 2017 E
% Last Change: Wed Oct 11 11:00 PM 2017 E
%
\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{tikz}

\pagestyle{fancy}

\lhead{CS 281}
\rhead{Final Project Abstract}

\begin{document}

\title{Final Project Proposal}
\author{
Kevin Chen              \\ \texttt{jiafengchen@college.harvard.edu}
\and Yufeng Ling        \\ \texttt{yufengling@college.harvard.edu}
\and Francisco Rivera   \\ \texttt{frivera@college.harvard.edu}}
\date{\today}

\maketitle

\begin{abstract}
Here.
\end{abstract}

\section{Requirements.txt}

\begin{enumerate}
\item Abstract
\item Changes since proposal
\item Baseline results
\end{enumerate}

\end{document}